---
title: "Final Project CIS 541"
class: "Fundaments of Big Data Analytics"
author: Ana Pacella
professor: Moises Vasquez
output: html_notebook
---

#Data exploration
```{r}
#summarizing the data
summary(wine_testing_data_curated)
head(wine_testing_data_curated)
```

```{r}
#taking only columns that we'll need
WineData <- wine_testing_data_curated[-c(1,5,6,7,9,10)]
WineData
```

#Visualizing the data
```{r}
ggplot(data = WineData, mapping = aes(x = points, y = price, color = country)) + 
  geom_point() + 
  labs(title = "Wine Data", subtitle = "Price vs Points", x = "Points", y = "Price")
```

#Grouping different types of wine and averaging its price and points
```{r}
WineTypes <- group_by(WineData, variety)
WineTypesSummary <- summarise(WineTypes, 
          avgPrice = mean(price),
          avgPoints = mean(points))
WineTypesSummary
```

#Comparing avg price of specific wine types: Pinot Noir and Chardonnay
```{r}
#selecting only Pinot Noir
PinotNoir <- subset(WineData, variety == "Pinot Noir")
PinotNoir
```


```{r}
#selecting only Chardonnay
Chardonnay <- subset(WineData, variety == "Chardonnay")
Chardonnay
```
# peroforming test statistics
```{r}
#performing wilcoxon test
wilcox.test(PinotNoir$price, Chardonnay$price)
```

#H0: There is no significant statistical difference between the average price of Pinot Noir and Chardonnay wines. 
#H1: There is significant statistical difference between the average price of Pinot Noir and Chardonnay wines.

#Since p-value is 2.2e-16, meaning that is less than 0.05, it means that there is a significant difference. There is enough evidence to reject H0 and support the claim.


```{r}
PinotChardonnay <- subset(WineData, variety == "Pinot Noir" | variety == "Chardonnay")
print(PinotChardonnay)
```


```{r}
ggplot(data = PinotChardonnay, mapping = aes(x = points, y = price, color = variety)) + 
  geom_point() + 
  geom_smooth(method = 'loess') + 
  labs(title = "Wine Types: Chardonnay and Pinot Noir", subtitle = "Price vs Points", x = "Points", y = "Price")
```

#Now, proceeding to perform regression analysis
```{r}
#performing regression analysis by separating data into training and testing sets

#setting random seed for repeatability
set.seed(123)

#splitting the data
sample <- sample.split(wine_testing_data_curated$points, SplitRatio = .75)
train <- subset(wine_testing_data_curated, sample == TRUE)
test <- subset(wine_testing_data_curated, sample == FALSE)
```


```{r}
#polynomial regression model
loessmodel <- loess(points ~ price, data = train)
loessmodel
```

```{r}
#summarizing results of loess function
summary(loessmodel)
```

```{r}
#predicting results of loess function
predictloess <- predict(loessmodel, test)
predictloess
```

```{r}
#presenting prediction in data frame to make it more legible
loesspredictionchart <- data.frame(variety = wine_testing_data_curated$variety, points = wine_testing_data_curated$points, predictloess)
loesspredictionchart
```

```{r}
#loess chart with averages  
PredictionChartLoessVar <- group_by(loesspredictionchart, variety)
PredictionChartLoessVar <- summarise(PredictionChartLoessVar, 
          avgPoints = mean(points),
          avgPredictedPrice = mean(predictloess))
PredictionChartLoessVar
```

```{r}
#checking error on loess
loessrmse <- rmse(wine_testing_data_curated$price,predictloess)
loessrmse
```


```{r}
#linear regression model
lmmodelreal <- loess(points ~ price, data = train)
lmmodelreal
```

```{r}
#summarizing lineal regression
summary(lmmodelreal)
```

```{r}
#predicting lineal regression
predictlm <- predict(lmmodel, test)
predictlm
```

```{r}
#presenting prediction in data frame to make it more legible
lmpredictionchart <- data.frame(variety = wine_testing_data_curated$variety, points = wine_testing_data_curated$points, predictlm)
lmpredictionchart
```


```{r}
#lm chart with averages  
PredictionChartLmVar <- group_by(lmpredictionchart, variety)
PredictionChartLmVar <- summarise(PredictionChartLmVar, 
          avgPoints = mean(points),
          avgPredictedPrice = mean(predictlm))
PredictionChartLmVar
```

```{r}
#checking error on lm
lmrmse <- rmse(wine_testing_data_curated$price,predictlm)
lmrmse
```

#performing regression analysis only with lm and loess 
```{r}
#linear regression model
lmmodel1 <- lm(points ~ price, data = WineTypes)
lmmodel1
```

```{r}
#summarizing lm
summary(lmmodel1)
```

```{r}
#predicting lm
predictlm1 <- predict(lmmodel1, newdata = wine_testing_data_curated)
predictlm1
```


```{r}
#Framing lm data to have it more organized
predictlmchart1 <- data.frame(variety = wine_testing_data_curated$variety, points = wine_testing_data_curated$points, predictlm1)
predictlmchart1
```


```{r}
#Checking error on lm
require(Metrics)
lmrmse1 <- rmse(wine_testing_data_curated$points,predictlm1)
lmrmse1
```

```{r}
#nonlinear regression model 
loessmodel1 <- loess(points ~ price, data = WineTypes)
loessmodel1
```


```{r}
#summarizing nonlinear regression model
summary(loessmodel1)
```


```{r}
#predicting with loess
predictloess1 <- predict(loessmodel1, newdata = wine_testing_data_curated)
predictloess1
```

```{r}
#Framing loess data to have it more organized
predictloesschart1 <- data.frame(variety = wine_testing_data_curated$variety, points = wine_testing_data_curated$points, predictloess1)
predictloesschart1
```

```{r}
require(tidyverse)
ggplot(data = predictloesschart1, mapping = aes(x = points, y = predictloess1)) +
  geom_point(color = 'brown') + 
  labs(title = "Price Predicted for Different Types of Wine", subtitle = "Price Predicted from Points", x = "Points", y = "Price Predicted")
```


```{r}
#checking loess error
loessrmse1 <- rmse(wine_testing_data_curated$points,predictloess1)
loessrmse1
```

#After conducting RMSE for each model, we can conclude that the model that best suits this analysis is the polynomial regression model (loess), since its error is less (2.53 vs 2.78).

```{r}
install.packages("dplyr")

```

```{r}
WineGrouped <- group_by(PinotChardonnay,variety)
summarise(PinotChardonnay,
          avgPrice = mean(price),
          avgPrice = mean(price))
```

